Warming non-randomly speeds up food web fluxes in both an absolute and relative sense

### 2022-02-14 doc/ms/ms.Rmd
finalizing spp rel flux figure w/ taxonomy. First draft of introduction

---
title: "Manuscript title here"
author:
  - James R. Junker:
      email: jrjunker@mtu.edu
      institute: [msu, glrc]
      correspondence: true
  - name: Wyatt F. Cross
    institute: msu
  - name: James M. Hood
    institute: osu
  - name: Jonathan P. Benstead
    institute: bama
  - name: Alex D. Huryn
    institute: bama
  - name: Daniel Nelson
    institute: usu
  - name: Jon S. Olafsson
    institute: veidi
  - name: Gisli M. Gislason
    institute: reyk
institute:
  - msu: Dept. of Ecology, Montana State University, Bozeman, MT USA
  - glrc: Current `address:` Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - osu: 
  - bama: 
  - usu: 
  - veidi: 
  - reyk: 
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

Energy fluxes through invertebrate food webs restructured in response to temperature. Specifically, communities in warmer streams were composed of smaller-bodied and higher-turnover populations on average. Additionally, energy fluxes of warmer streams were unequally skewed towards small-bodied, high-turnover populations *within* the community, suggesting higher temperatures restructured energy fluxes in both an absolute and relative sense across and within communities.

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures are altering the provisioning and maintenance of ecosystem services by modifying the network of interactions among species that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. Across global climate gradients, temperature mediates the dynamics and stability of food webs [@baiser2019] by altering the acquisition and allocation of resources among species [@zhang2017] and ultimately the magnitudes and relative distribution of biomass and interactions among species and trophic levels [@may1972; @mccann1998; @barnes2018; @gibert2019]. Temperature's effects permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001, @brown2004] and biological activity [e.g., attack rate, handling time, growth rates, etc.; @dell2014], as an environmental niche filter of community assembly [] and indirectly through the distribution of species traits, such as body size, within and across species [@bideault2019; @nelson2017].

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
In fact, reduced organismal body size with warming is considered a 'universal' response to climate change [@atkinson1994; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, such as the increased relative abundance of smaller species within warmer communities [@bergmann1848], or smaller individuals in warmer populations [@james1970], or lastly through reduced body size of warmer individuals [@atkinson1994]. Temperature-size relationships (TSR) are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes because body size is a strong determinant of organismal biology [@peters1983], influencing life history [@zuess2017; @altermatt2010] and development [@angilleta], diet breadth [], trophic position [], and crucially, metabolic rate [@brown2004].

Organismal metabolic rate, in addition to being controlled by body size, is strongly influenced by  temperature [@gilooly2001; @arrhenius1898]. Therefore, temperature may alter metabolic rate indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019]. The interactive effect of these processes has been shown to modulate ecosystem patterns through effects on population abundance [@bernhardt2019], consumer-resource interactions [@bideault20149], and food web structure [@gibert2019]. Importantly, the kinetic effects of temperatures are tied to many biological processes [@dell2014], among them growth rate [@giloolly2001`], developmental rate [@zuo2012], and population turnover rate [@]. As such, the emerging picture suggests a smaller, faster world in which warming can "speeds up" ecosystem processes through the compounded effects of smaller body size and higher biological rates.

Yet, considerable uncertainty remains in the predicted short- and long-term effects of climate warming, especially at higher levels of organization such as communities and food webs *more*[@woodward2010; @walther2005]. For example, we lack a complete understanding of TSRs to explain deviations that exist across taxonomic groups [e.g., diatoms, @adams2013; invertebrates @zeuss2017; birds @geist1987; and fish @rypel2014]. Explanations for these deviations vary [@ohlberger2013], some highlight additional processes such as competition [] and predation [@delong2014], while others propose alternative models of organism body size optimization, such as the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012]. Crucially, these alternative models highlight the trade-off between asymptotic body size and metabolic demand [@delong2012] that can shape life history patterns [i.e., voltinism, @zuess2017].  and community structure across ecosystems...*better transition here maybe*

These maxmize energy intake when resources are available [@atkinson1997]

- Seasonal systems have unique constraints
- These seasonal constraints are greater at higher temps
- This creates a situation where body sizes are selected for that 1) stabilize c-r dynamics [@osmond2017],
2) availability

Here, we measured the patterning and distribution of energy fluxes within invertebrate food webs across a natural stream temperature gradient (\~5--28$^\circ$C). Previous research in these streams has shown a strong positive effect of temperature on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally
[@ogorman2012; @hood2018]. Consumers rely largely on autochthonous resources [@ogorman2012; @nelson2020] and therefore the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. As such, we predicted annual energy fluxes of consumers to scale with among-stream patterns in resource availability and consumer energy demand and increase with temperature across streams. We hypothesized temperature to be a strong environmental filter of community assembly and energy flux, especially at the extremes. From this we propose a number of predictions regarding the patterning of energy fluxes within and across communities of different temperature, with particular focus on two species traits: body size and population biomass turnover (production:biomass ratios, *P:B*). First, we predict warmer temperatures should select for smaller and faster turnover species leading to a decrease in body sizes and increase in *P:B* on average across communities. Second, within communities energy fluxes should be dominated by smaller, higher turnover species at warmer temperatures. Third,

e further expected that increasing temperature would reduce consumer species richness [@ogorman2019], thereby altering how OM fluxes are distributed within and across communities. We predicted OM fluxes would be most even at moderate temperatures and dominated by few taxa at the coldest and warmest extremes. Lastly, we predicted that temperature is a strong filter of community assembly  Lastly, we quantified how OM fluxes were associated with two key species traits; body size and population production to biomass ratio (P:B) within communities. We predicted mean community body size would decrease and mean community P:B ratios would increase at higher temperatures. The distribution of OM fluxes to consumers will shift towards relatively faster, higher turnover consumer at higher temperatures, 'speeding up' consumer dynamics both through direct effects on consumer turnover and through a decrease in mean body size.

@@

Macroinvertebrate diets were quantified for dominant taxa in each stream. We focused on numerically abundant taxa and/or taxa with relatively
high annual production. A minimum of five individuals were selected from samples, and, when possible included individuals of different size classes to account for ontogenetic
shifts in diet. We included individuals from different seasons to capture concurrent ontogenetic and seasonal changes. For small-bodied taxa, we combined multiple individuals (*n*
= 3--5) to ensure samples contained enough material for quantification. We used methods outlined in Rosi-Marshall [-@rosi-marshall2016] to remove gut tracts and prepare gut contents for quantification. Briefly, we removed the foregut from each individual or collection of individuals and sonicated contents in water for 30 seconds. Gut content slurries were filtered
onto gridded nitrocellulose membrane filters (Metricel GN-6, 25 mm, 0.45 $\mu$m pore size; Gelman Sciences, Ann Arbor, MI, USA), dried at 60 $^\circ$C for 15 min, placed on a
microscope slide, cleared with Type B immersion oil, and covered with a cover slip. We took 5--10 random photographs under 200--400x magnification, depending on the density of
particles, using a digital camera mounted on a compound microscope. From these photographs we identified all particles within each field and measured the relative area of particles
using image analysis software [@schindelin2012]. We identified particles as diatoms, green and filamentous algae, cyanobacteria, amorphous detritus, vascular and non-vascular
plants (e.g., bryophytes), and animal material. We calculated the proportion of each food category in the gut by dividing their summed area by the total area of all particles. Gut
contents of many predators were empty or contained unidentifiable, macerated prey. For these taxa, we assumed 100% animal material.

To estimate variability in diet compositions and to impute missing values for non-dominant, yet present, taxa, we modeled the diet proportions within each stream using a
hierarchical multivariate model [@fordyce2011, @coblentz2017]. Here, the diet of a consumer population, *i*, in stream, *j*, is a multinomial vector,$\overrightarrow{y_{ij}}$, of
$$\overrightarrow{y_{ij}} \sim Multinomial(\overrightarrow{p_{ij}}, n_{ij})$$ $$\overrightarrow{p_i} \sim Dirichlet(\overrightarrow{q_i} \times  \alpha) $$ where, $\overrightarrow{p_i}$, is
a vector of consumer diet proportions, $\overrightarrow{q_i}$ is a vector of the population's diet proportions and $\alpha$ is a concentration parameter of the Dirichlet process.
We used uniform priors for $\overrightarrow{q_i}$ and $\alpha$, $$\overrightarrow{q_i} \sim Dirichlet(\overrightarrow{\textbf{1}})$$ $$\alpha \sim Uniform(0,\textit{c})$$ where,
$\overrightarrow{\textbf{1}}$ is a vector of ones the same length of basal resource types and $\textit{c}$ is the presumed maximum value the concentration value. Models were fit in Stan with the 'brms' package in R [@burkner2017]. For non-dominant taxa, diet proportions were imputed
from the hierarchical model by resampling from posterior distributions. Importantly, this process allowed to maintain the hierarchical structure of the data when imputing missing values.

From modeled diet compositions, we estimated trophic redundancy within and across stream food webs by calculating proportional similarities [*PS*; @whittaker1952] among modeled diet estimates. Proportional similarities were calculated as:

@@

To visualize and quantify how evenly OM fluxes were distributed among consumers within a stream, we constructed Lorenz curves [@lorenz1905] on ordered relative consumption fluxes, such that in a community with $S$ species and the relative consumption of species *i*, $p_i$, is ordered $p_1 \leq p_2 \leq ... p_S$. The Lorenz curve plots how a value, in this case OM flux, accumulates with increasing cumulative proportion of species. In a community with perfectly equal distribution of OM consumption among species, the Lorenz curve is simply a straight diagonal line. Deviation from perfect equality was calculated as the Gini coefficient [@gini1921], normalized for differences in $S$ among streams, $G^*$ [@solomon1975; @chao2019]: $$ G^* = (2 \sum_{i = 1}^S ip_i -2)/(S-1)$$

### Distribution along species' trait axes

We were interested in the question of how temperature may select for certain species traits (i.e., body size, $P:B$ ratio, population biomass) and therefore OM fluxes would skew towards populations with those traits. To detect this, for all bootstrapped estimates of annual OM flux, we ordered species based on within-stream ranking of annual population traits (i.e, $M$, $P:B$) and then calculated a measure of skewness, $Sk_{flux}$, based on quartiles of the distribution of OM fluxes in relation to species traits as:

@@

where, $f_{Qx}$, is the cumulative flux at some quantile, $Qx$, of the community trait distribution. To determine if this outcome was due to random ordering, we first had to account for statistical constraints that restrict the range of possible outcomes [\ i.e., feasible set, @haegeman2008; @diaz2021a], given the number of species and the relative distribution of energy fluxes within a community. The number of unique orderings of species increases to computationally intractable numbers very quickly (e.g., $S!$, 10 species ~3.6e^6^ unique orderings). Therefore, we chose to permute a portion of each stream community's feasible set by randomly ordering species and calculating the skewness in the cumulative distribution of annual OM fluxes 100,000 times. The number of random orderings was chosen as a balance between characterizing the relative distribution of skewness values within each feasible set and computational and time constrains. This permuted set allowed us to calculate the probability of observing the empirical skewness, $Sk_{flux}$, in each stream compared to a random ordering given the distribution of relative energy flux.

## Statistical Analyses

Relationships between mean annual temperature ($^\circ$C) and mean community *M* and *P:B* were assessed with bootstrapped linear regressions. Here, `r temperature_stats[['n_boot']]` values of mean *M* or *P:B* were resampled with replacement from each stream. For each resampling event, a linear model was fit between *log~e~*-transformed *M* or *P:B* and mean annual temperature. Response variables *M* and *P:B* were transformed to meet the assumption of normally distributed residual variation. We similarly used a bootstrapped beta regression approach to test for relationships between annual stream temperature and the skewness of fluxes, $Sk_{flux}$, with *M* and *P:B*. Skewness coefficients exist in the range [-1, 1], therefore we applied a simple transformation, 
$$ (Sk_{flux} + 1)/2 $$ 
to standardize the values between 0 and 1 to meet the assumptions of the model. Coefficients were back-transformed to estimate effect sizes. To detect changes in 'random' vs. non-random assembly, we  again used beta regression on estimates of the probability of seeing the empirical $Sk_{flux}$ value or one more extreme based on random ordering of species in the permuted feasible set. Here, values can take the range between 0 and 1 so no transformation was necessary. As necessary, when we observed many values near 0 and 1, we implemented a zero-one inflated beta regressions in brms [@burkner2017] to test if the probability of observing the empirical $Sk_{flux}$ was associated with mean annual temperature across streams.

@@

@@

Annual community energy demand mirrored patterns of secondary production previously reported [@junker2020]. Community energy demand varied ~`r round((max_demand$flux_mg_m_y_mean/min_demand$flux_mg_m_y_mean),1)`-fold across streams (`r round((min_demand$flux_mg_m_y_mean/1000), 3)`; `r round((min_demand$flux_mg_m_y_quant2.5/1000),3)` -- `r round((min_demand$flux_mg_m_y_quant97.5/1000),3)` to `r round((max_demand$flux_mg_m_y_mean/1000),1)`; `r round((max_demand$flux_mg_m_y_quant2.5/1000),1)` -- `r round((max_demand$flux_mg_m_y_quant97.5/1000),1)` g AFDM $m^{-2} y^{-1}$ mean; 95% percentile interval (PI)) and was positively related to temperature. Total OM flux through the consumer community was strongly associated with community energy demand and varied from `r round((min_flux$flux_mg_m_y_mean/1000),1)`; `r round((min_flux$flux_mg_m_y_quant2.5/1000),1)`--`r round((min_flux$flux_mg_m_y_quant97.5/1000),1)` to `r round((max_flux$flux_mg_m_y_mean/1000),1)`; `r round((max_flux$flux_mg_m_y_quant2.5/1000),1)`--`r round((max_flux$flux_mg_m_y_quant97.5/1000),1)`
g AFDM $m^{-2} y^{-1}$.

General patterns of community change and species identity across the gradient.
Community and population patterns of M and P:B with temperature...

@@

**I am strongly considering ditching or appendixing the diet stuff. This will become important for the next paper, however, I think it distracts from the story here since we are really just focused on the energy pathways**

@@

Variation in the relative distribution of OM fluxes among consumers corresponded to different patterns of dominance across streams (Figure 1). For example, in an absolute sense, ~85% of total flux was contributed by `r n_dom_min$n_dom` to `r n_dom_max$n_dom` species. Relatively, this flux was attributed by `r round(perc_dom_min$dom_spp*100,0)`% to `r round(perc_dom_max$dom_spp*100,0)`% of the species assemblages within streams. Differences in dominance were unrelated to temperature, however, species dominance in relation to population traits exhibited important differences along the temperature gradient.

```{r figure 1., echo = FALSE, warning = FALSE, message = FALSE, fig.width=5, fig.height=5, dpi = 450, fig.cap="Figure 1. Rank ordering of annual population energy flux log~e~(mg m^-2^ y^-1^) in consumer communities across the temperature gradient."}
drake::loadd(annual_spp_flux_fig)
annual_spp_flux_fig
```

@@

>>>>

---
title: "Warming non-randomly speeds up food web fluxes in both an absolute and relative sense"
author:
  - James R. Junker:
      email: jrjunker@mtu.edu
      institute: [msu, glrc]
      correspondence: true
  - name: Wyatt F. Cross
    institute: msu
  - name: James M. Hood
    institute: osu
  - name: Jonathan P. Benstead
    institute: bama
  - name: Alex D. Huryn
    institute: bama
  - name: Daniel Nelson
    institute: usu
  - name: Jon S. Olafsson
    institute: veidi
  - name: Gisli M. Gislason
    institute: reyk
institute:
  - msu: Department of Ecology, Montana State University, Bozeman, MT USA
  - glrc: Current `address:` Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - osu: Department of Evolution, Ecology, and Organismal Biology, The Ohio State University, Translational Data Analytics Institute, The Aquatic Ecology Laboratory, Columbus, OH 43212, USA
  - bama: Department of Biological Sciences, University of Alabama, Tuscaloosa, AL  35487, USA
  - usu: National Aquatic Monitoring Center, Department of Watershed Sciences, Utah State University, Logan, UT 84322, USA
  - veidi: Institute of Marine and Freshwater Fisheries, Reykjavik, Iceland
  - reyk: Univeristy of Iceland, Institute of Life and Environmental Sciences, Reykjavik, Iceland
output:
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   pdf_document:
     keep_tex: true
   html_document: default
   pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \usepackage{indentfirst}
- \linenumbers
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

Warming temperatures are altering biological communities and trophic networks globally. While the influence of warming in food webs is often context-dependent, increasing temperatures are predicted to alter to alter food webs and ecosystems in a couple fundamental ways by 1) decreasing organism body size and 2) increasing metabolic rates. These two attributes, body size and metabolic rate, are important drivers of many ecological patterns. Therefore, warming-induced changes have the potential to propagate through countless ecosystem processes thereby altering the distribution of food webs fluxes, food web stability, and potentially the importance of environmental and biological factors in community assembly. Here, we quantify the patterning and relative distribution of organic matter fluxes through stream food webs spanning a ~25$^\circ$C temperature gradient. We then relate patterns in fluxes to species and community traits (body size, *M* and biomass turnover, *P:B*) within and across ecosystems. We predicted 1) fluxes would skew towards smaller *M* and higher *P:B* both across and within ecosystems and 2) fluxes would be more skewed at higher temperatures. Across the temperature gradient, communities in warmer streams were composed of smaller-bodied and higher-turnover populations on average. Additionally, material fluxes of warmer streams were increasingly skewed towards high-turnover populations *within* the community, showing higher temperatures restructured organic matter fluxes in both an absolute and relative sense. Lastly, in warmer ecosystems the relative distribution of organic matter fluxes appears to be increasingly 'non-random', evidence of potentially stronger selection for these traits with increasing temperature. The emerging picture is a warmer world that is both smaller and faster. Our study lends further support to this pattern and also suggests that

@@

<!--- warming and effects in food webs --->
<!--- Speeding up of food web: Species levels--->
Increasing global temperatures are altering the provisioning and maintenance of ecosystem services by modifying the network of interactions among species that underpin ecosystem functions [@brose2012; @woodward2010; @deruiter1995; @thompson2012]. Across global climate gradients, temperature mediates the dynamics and stability of food webs [@baiser2019] by altering the acquisition and allocation of resources among species [@zhang2017] and ultimately the magnitudes and relative distribution of biomass and fluxes among species and trophic levels [@may1972; @mccann1998; @barnes2018; @gibert2019a]. Temperature's effects permeate across levels of biological organization from its control on individual metabolic rates [@gillooly2001; @brown2004] and biological activity (e.g., attack rate, handling time, growth rates, etc.; @dell2014, as an environmental niche filter of community assembly and indirectly through the distribution of species traits, such as body size, within and across species [@bideault2019; @nelson2017].

<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
In fact, reduced organismal body size with warming is considered a 'universal' response to climate change [@atkinson1994; @gardner2011; @daufresne2009]. Body size reductions arise from multiple mechanisms acting at different levels of organization, such as the increased relative abundance of smaller species within warmer communities [@bergmann1848], or smaller individuals in warmer populations [@james1970], or through reduced body size of warmer individuals [@atkinson1994]. Temperature-size relationships (TSR) are likely to influence the distribution of body sizes both across and within ecosystems, such that warmer communities should contain smaller species and populations, on average, and smaller organisms should be more dominant within a community. These changes can have important implications for ecosystems pattern and processes because body size is a strong determinant of an organism's biology [@peters1983] influencing, life history [@zeuss2017; @altermatt2010; @nelson2020b], development [@angilletta2004], and crucially, basal metabolic rate [@brown2004; @gillooly2001].

<!--- Write the main goals of the paragraph--->
Organismal metabolic rate, in addition to being controlled by body size, is strongly influenced by temperature [@gillooly2001]. Therefore, temperature may alter metabolic rate indirectly through reductions in body size, as well as directly through its effects on subcellular kinetics [@bideault2019; @osmond2017]. The interactive effect of these processes has been shown to modulate ecosystem patterns through effects on population abundance [@bernhardt2018], consumer-resource interactions [@bideault2019], and food web structure [@gibert2019a]. Importantly, the kinetic effects of temperatures are tied to many biological processes [@dell2014], among them growth rate [@gillooly2001], developmental rate [@zuo2012; @nelson2020b], voltinism [@zeuss2017], and population turnover rate [@huryn2007; @brown2004]. As such, the emerging picture suggests a smaller, faster world in which warming "speeds up" ecosystem processes through the compounded effects of smaller body size and higher turnover rates.

Yet, considerable uncertainty remains in the predicted short- and long-term effects of climate warming, especially at higher levels of organization such as communities and food webs [@woodward2010; @walther2002]. For example, species interactions in food webs may modulate species' TSRs [@delong2014], leading to varied responses among species and trophic levels [@ohlberger2013]. Additionally, warming-induced body size reductions may act as a stabilizing process in some consumer-resource interactions through reduced consumer:resource biomass ratios [@osmond2017] and reductions in consumer growth rates relative to resource supply [@kozlowski2004; @delong2012a; @mccann2011; @greyson-gaito2022]. 
Therefore, food web processes may impart varying selective pressures on the distribution of body sizes and population turnover within communities. If strong, these pressures may leave distinct imprints on community structure and the distribution of energy fluxes within food webs with potential implications for the stability and functioning of ecosystems on a warming planet [@petchey1999; @fussmann2014].

Here, we measured the patterning and distribution of energy fluxes within invertebrate food webs across a natural stream temperature gradient (\~5--28$^\circ$C). Previous research in these streams has shown a strong positive effect of temperature on primary production both among streams [@demars2011; @padfield2017] and within streams seasonally
[@ogorman2012; @hood2018]. Consumers rely largely on autochthonous resources [@ogorman2012; @nelson2020] and therefore the dynamics of primary production have a strong control on consumer energy demand [@junker2020]. As such, we predicted annual energy fluxes of consumers to scale with among-stream patterns in resource availability and consumer energy demand and increase with temperature across streams. We hypothesized temperature to be a strong environmental filter of community assembly and the distribution of energy and organic matter fluxes, especially at the extremes. From this we made a number of predictions regarding the patterning of energy fluxes within and across communities of different temperature, with particular focus on two species traits: body size and population biomass turnover (production:biomass ratios, *P:B*). First, we predict warmer temperatures should favor smaller and faster turnover species leading to a decrease in body sizes and increase in *P:B* on average across communities. Second, within communities energy fluxes should be dominated by smaller, higher turnover species at warmer temperatures. Third, we hypothesized that the selective pressure from environmental temperature on body sizes and population turnover would be greatest at temperature extremes, therefore, the patterning of energy fluxes will be 'non-random' at coldest and warmest temperatures.

@@

Macroinvertebrate diets were quantified for dominant taxa in each stream. We focused on numerically abundant taxa and/or taxa with relatively
high annual production. A minimum of five individuals were selected from samples, and, when possible included individuals of different size classes to account for ontogenetic
shifts in diet. We included individuals from different seasons to capture concurrent ontogenetic and seasonal changes. For small-bodied taxa, we combined multiple individuals (*n*
= 3--5) to ensure samples contained enough material for quantification. We used methods outlined in Rosi-Marshall [-@rosi-marshall2016] to remove gut tracts and prepare gut contents for quantification. Briefly, we removed the foregut from each individual or collection of individuals and sonicated contents in water for 30 seconds. Gut content slurries were filtered
onto gridded nitrocellulose membrane filters (Metricel GN-6, 25 mm, 0.45 $\mu$m pore size; Gelman Sciences, Ann Arbor, MI, USA), dried at 60 $^\circ$C for 15 min, placed on a
microscope slide, cleared with Type B immersion oil, and covered with a cover slip. We took 5--10 random photographs under 200--400x magnification, depending on the density of
particles, using a digital camera mounted on a compound microscope. From these photographs we identified all particles within each field and measured the relative area of particles
using image analysis software [@schindelin2012]. We classified particles into six categories: diatoms, green and filamentous algae, cyanobacteria, amorphous detritus, vascular and non-vascular
plants (e.g., bryophytes), and animal material. We calculated the proportion of each food category in the gut by dividing their summed area by the total area of all particles. Gut
contents of many predators were empty or contained unidentifiable, macerated prey. For these taxa, we assumed 100% animal material.

To estimate variability in diet compositions and to impute missing values for non-dominant, yet present, taxa, we modeled the diet proportions within each stream using a
hierarchical multivariate model [@fordyce2011; @coblentz2017]. Here, the diet of a consumer population, *i*, in stream, *j*, is a multinomial vector,$\overrightarrow{y_{ij}}$, of
$$\overrightarrow{y_{ij}} \sim Multinomial(\overrightarrow{p_{ij}}, n_{ij})$$ $$\overrightarrow{p_i} \sim Dirichlet(\overrightarrow{q_i} \times  \alpha) $$ where, $\overrightarrow{p_i}$, is
a vector of consumer diet proportions, $\overrightarrow{q_i}$ is a vector of the population's diet proportions and $\alpha$ is a concentration parameter of the Dirichlet process.
We used uniform priors for $\overrightarrow{q_i}$ and $\alpha$, $$\overrightarrow{q_i} \sim Dirichlet(\overrightarrow{\textbf{1}})$$ $$\alpha \sim Uniform(0,\textit{c})$$ where,
$\overrightarrow{\textbf{1}}$ is a vector of ones the same length of basal resource types and $\textit{c}$ is the assumed concentration value. Models were fit in Stan with the 'brms' package in R [@burkner2017]. For non-dominant taxa, diet proportions were imputed
from the hierarchical model by resampling from posterior distributions. Importantly, this process allowed to maintain the hierarchical structure of the data when imputing missing values.

From modeled diet compositions, we estimated trophic redundancy within and across stream food webs by calculating proportional similarities (*PS*; @whittaker1952) among modeled diet estimates. Proportional similarities were calculated as:

@@

To visualize and quantify how evenly OM fluxes were distributed among consumers within a stream, we constructed Lorenz curves [@lorenz1905] on ordered relative consumption fluxes, such that in a community with $S$ species and the relative consumption of species *i*, $p_i$, is ordered $p_1 \leq p_2 \leq ... p_S$. The Lorenz curve plots how a value, in this case OM flux, accumulates with increasing cumulative proportion of species. In a community with perfectly equal distribution of OM consumption among species, the Lorenz curve is simply a straight diagonal line. Deviation from perfect equality was calculated as the Gini coefficient [@gini1921], normalized for differences in $S$ among streams, $G^*$ [@solomon1975; @chao2019]:

$$ G^* = (2 \sum_{i = 1}^S ip_i -2)/(S-1)$$

where $G^*$ represents an index of relative evenness of OM fluxes bounded between zero and one--one representing a community with exactly equal proportion of total community OM flux for all species ($1/S$), and a value of zero where total community flux is attributed to a single species.

### Distribution of OM fluxes in relation to species' traits

We predicted that warming would favor species with smaller body size and higher population turnover and therefore OM fluxes would be skewed towards small body size (*M*) and higher *P:B* across and within communities. Across communities, we assessed the relationships between mean annual temperature ($^\circ$C) and mean population *M* and *P:B* of each community with bootstrapped linear regressions. Here, `r temperature_stats[['n_boot']]` values of population *M* or *P:B* were resampled with replacement for each species within each stream. The mean of all populations within a stream was calculated and a linear model was fit between *log~e~*-transformed *M* or *P:B* and mean annual temperature. Response variables *M* and *P:B* were transformed to meet the assumption of normally distributed residual variation.

To quantify the distribution of energy fluxes _**within**_ a community, we assessed if relative OM flux was skewed towards populations with lower or higher relative *M* or  *P:B*. To do this, we ordered species based on within-stream ranking of annual population traits (i.e, $M$, $P:B$) and then calculated a measure of skewness, $Sk_{flux}$, based on quartiles of the distribution of OM fluxes in relation to species traits as:

@@

where, $f_{Qx}$, is the cumulative flux at some quantile, $Qx$, of the community trait distribution. We repeated this analysis for all bootstrapped estimates of OM flux in all communities. Skewness coefficients exist in the range [-1, 1], where -1 indicates OM fluxes are skewed perfectly away from a trait and 1 indicates higher relative flux is perfectly associated with higher trait values. To determine if $Sk_{flux}$ with population *M* and *P:B* was related to mean annual stream temperature, we use bootstrapped beta regression with a simple transformation to meet the assumptions of the model, $(Sk_{flux} + 1)/2$, thereby standardizing values between 0 and 1. Model coefficients were back-transformed to estimate effect sizes.

To assess the strength of environmental filtering of species' traits within a community, we quantified how likely it was to observe such a skewed distribution in OM fluxes by random chance. Here, we predicted that communities with strong filtering for traits would exhibit highly skewed OM distributions that would be unlikely due to chance (i.e., 'non-random ordering'). In contrast, weak environmental filtering would produce OM fluxes with skew values that are likely due to random chance ('random ordering'. To accomplish this, we first had to account for statistical constraints that restrict the range of possible outcomes [\ i.e., feasible set, @haegeman2008; @diaz2021a], given the number of species and the relative distribution of OM fluxes within a community. The number of unique orderings of species increases to computationally intractable numbers very quickly (e.g., $S!$, 10 species ~3.6e^6^ unique orderings). Therefore, we chose to permute a random subset of each stream community's feasible set by randomly ordering species and calculating the skewness in the cumulative distribution of annual OM fluxes 100,000 times in each stream. The number of random orderings was chosen as a balance between characterizing the distribution of skewness values within each feasible set and computational and time constrains. This permuted set allowed us to calculate the probability of observing the empirical skewness, $Sk_{flux}$, in each stream compared to a random ordering given the distribution of relative OM flux. To assess the relationship between the probability of 'non-random ordering' and temperature, we again used beta regression on estimates of the probability of seeing the empirical $Sk_{flux}$ value or one more extreme based on random ordering of species in the permuted feasible set. Here, values can take the range between 0 and 1 so no transformation was necessary. As necessary, when we observed many values near 0 and 1, we implemented a zero-one inflated beta regressions in brms [@burkner2017] to test if the probability of observing the empirical $Sk_{flux}$ was associated with mean annual temperature across streams. We predicted increasingly 'non-random' distributions in OM fluxes with increasing temperature.

@@

pb_spp_temp_coefs = temperature_stats[["pb_spp_temp_coefs"]]
m_spp_temp_coefs = temperature_stats[["m_spp_temp_coefs"]]

@@

Annual community energy demand mirrored patterns of secondary production previously reported [@junker2020]. Community energy demand varied ~`r round((max_demand$flux_mg_m_y_mean/min_demand$flux_mg_m_y_mean),1)`-fold across streams (`r round((min_demand$flux_mg_m_y_mean/1000), 3)`; `r round((min_demand$flux_mg_m_y_quant2.5/1000),3)` -- `r round((min_demand$flux_mg_m_y_quant97.5/1000),3)` to `r round((max_demand$flux_mg_m_y_mean/1000),1)`; `r round((max_demand$flux_mg_m_y_quant2.5/1000),1)` -- `r round((max_demand$flux_mg_m_y_quant97.5/1000),1)` g AFDM $m^{-2} y^{-1}$ mean; 95% percentile interval (PI)) and was positively related to temperature. Total OM flux through the consumer community was strongly associated with community energy demand and varied from `r round((min_flux$flux_mg_m_y_mean/1000),1)` (`r round((min_flux$flux_mg_m_y_quant2.5/1000),1)`--`r round((min_flux$flux_mg_m_y_quant97.5/1000),1)`, 95% PI) to `r round((max_flux$flux_mg_m_y_mean/1000),1)` (`r round((max_flux$flux_mg_m_y_quant2.5/1000),1)`--`r round((max_flux$flux_mg_m_y_quant97.5/1000),1)`, 95% PI) g AFDM $m^{-2} y^{-1}$.

@@

Generally, organic matter fluxes were dominated by insect species from *Simuliidae* and *Chironomidae* families (Figure 1a). In warmer streams, OM fluxes shifted toward the pulmonate snail, *Radix balthica*. In the warmest stream, where maximum temperatures approach ~40$^\circ$C, OM fluxes were dominated by worms of family *Naididae*, the midge *Cricotpous sylvestris*, and *Radix balthica*.

```{r figure 1, echo = FALSE, warning = FALSE, message = FALSE, fig.width=7.5, fig.height=6, dpi = 450, fig.cap="Figure 1. Patterns of OM flux among populations based on (a) Rank ordering of annual population organic matter flux log~e~(mg m^-2^ y^-1^) and (b) relative OM flux among taxaonomic groups in consumer communities and across the temperature gradient."}
drake::loadd(annual_spp_flux_fig)
drake::loadd(spp_flux_dist)
gridExtra::grid.arrange( spp_flux_dist,annual_spp_flux_fig, ncol = 2, layout_matrix = cbind(1,1,1,1,2,2,2))
```

```{r trait df, echo = FALSE, warning = FALSE, message = FALSE}
drake::loadd(spp_traitsVtemp)
ann_spp_summ = spp_traitsVtemp[['ann_spp_summ']]

@@

Variation in the relative distribution of OM fluxes among consumers corresponded to different patterns of dominance across streams (Figure 1b). For example, in an absolute sense, ~85% of total flux was contributed by `r n_dom_min$n_dom` to `r n_dom_max$n_dom` species. Relatively, this flux was attributed by `r round(perc_dom_min$dom_spp*100,0)`% to `r round(perc_dom_max$dom_spp*100,0)`% of the species assemblages within streams. Differences in dominance were unrelated to temperature, however, species dominance in relation to population traits exhibited important differences along the temperature gradient.

```{r trait v temp, echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=5, dpi = 450, fig.cap="Figure 2. The relationships between mean annual stream temperature and (a) average population body size of the community (mg^-ind^) and (b) average population turnover expressed as production to biomass ratios (*P:B*, y^-1^)."}

drake::loadd(spp_traitsVtemp)
grid.draw(spp_traitsVtemp[['fig1_plot']])

 # drake::loadd(spp_trait_histograms)
# grid.draw(spp_trait_histograms)

@@

Recent, models have proposed  However, 
- both bottom-up (e.g. resource supply) and top-down (e.g., predation) proceesses
- competition and predation [@delong2014], stabilize c-r dynamics [@osmond2017], optimization, such as the balance of resource supply and energy demand [e.g., @kozlowski2004; @delong2012a]
-
